---
title: "flight_delay"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flight_delay}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Predictive Modeling of Flight Delays Using `ridgereg()`

The purpose of this vignette is to showcase the use of the `ridgereg()` function to predict flight delays using the **nycflights13** dataset. This example demonstrates how to preprocess the data, incorporate weather features, and split the data for training, validation, and testing.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(heleramcar4)
library(nycflights13)
library(caret)
library(ggplot2)

set.seed(1)
```

## 1. Load and Clean the Dataset

We'll begin by loading the `flights` and `weather` datasets from the **nycflights13** package and remove variables that don't provide predictive value.

The `flights` dataset contains two delay variables: `dep_delay` (departure delay) and `arr_delay` (arrival delay). We'll use `dep_delay` as the target variable for our prediction since it's often the most relevant when discussing flight delays.

Next, we will discard variables that are not available when predicting departure delays, such as `arr_time`, `arr_delay`, `air_time`, and `dep_time`. Additionally, we remove variables that provide no predictive information:

-   **carrier:** Two-letter carrier abbreviation.
-   **flight:** Flight number.
-   **tailnum:** Plane tail number.

The following columns will be retained for modeling:

```{r}
# Define the subset of features from the flights dataset.
flights_features <- c(
  "dep_delay",  # Target variable
  "sched_dep_time", 
  "sched_arr_time", 
  "origin",
  "dest",
  "distance",
  "time_hour"
)
```

## 2. Merge with Weather Data

To improve prediction accuracy, we'll join the `weather` dataset with the `flights` dataset using the columns `origin` and `time_hour`. Only the selected features will be used in this merge.

```{r}
# Merge the weather data into the flights dataset.
X_merged <- merge(
  x = flights[flights_features], 
  y = weather, 
  by = c("origin", "time_hour"),  
  all.x = TRUE
)
```

Next, we'll remove columns with more than 10% missing values, and rows with any missing values will also be dropped.

```{r}
# Drop columns with more than 10% missing values, and remove rows with any missing data.
nan_count <- sapply(X_merged, function(y) sum(is.na(y)))
nan_pct <- (nan_count / nrow(X_merged)) * 100
drops <- names(nan_pct[nan_pct > 10])
X_nanless <- na.omit(X_merged[, !names(X_merged) %in% drops])
```

```{r echo=TRUE}
# Sanity check to ensure all missing values are removed.
nan_count <- sum(sapply(X_nanless, function(y) sum(is.na(y))))
sprintf("The total number of missing values is: %s", nan_count)
```

## 3. Split Data into Train, Validation, and Test Sets

Using the `caret` package, we'll split the cleaned dataset into training (80%), validation (15%), and test sets (5%).

We will only use the numerical columns from this point onward to simplify the modeling process.

```{r}
# Select only the numerical columns for modeling.
nums <- unlist(lapply(X_nanless, is.numeric), use.names = FALSE)  
X_num <- X_nanless[, nums]
```

We then create the train-validation-test split:

```{r}
# Split the dataset into training (80%), validation (15%), and test (5%) sets.

# First split: 80% for training, 20% for validation + test.
train_index <- createDataPartition(X_num$dep_delay, list = FALSE, p = .80, times = 1)
X_train <- X_num[train_index, ]
X_val_and_test <- X_num[-train_index, ]

# Second split: 75% of the remaining 20% for validation, 25% for test.
val_index <- createDataPartition(X_val_and_test$dep_delay, list = FALSE, p = .75, times = 1)
X_val <- X_val_and_test[val_index, ]
X_test <- X_val_and_test[-val_index, ]
```

## 4. Finding the Optimal Value of 位

The function `create_ridge_objective_function()` creates an objective function to evaluate the ridge regression model for different values of 位 (regularization parameter). The goal is to minimize the RMSE (Root Mean Squared Error) on the validation dataset.

```{r}
#' Create Ridge Objective Function
#'
#' This function generates an objective function that can be used to evaluate the 
#' performance of a ridge regression model on a validation set for a given value of lambda.
#'
#' @param formula The formula for the ridge regression model (e.g., dep_delay ~ .).
#' @param train The training dataset used to fit the ridge regression model.
#' @param val The validation dataset used to evaluate the model's performance.
#' @return A function that takes lambda as input and returns the RMSE (root mean squared error)
#'         for the predictions made by the ridge regression model on the validation data.
#'
#' @details The objective function returned by this function fits a ridge regression model
#'          using the training data and then predicts the target variable on the validation data.
#'          The model's performance is evaluated using RMSE.
create_ridge_objective_function <- function(formula, train, val) {
  
  # This inner function takes a lambda value, fits the ridge regression model,
  # predicts on the validation set, and returns the RMSE of the predictions.
  objective_function <- function(lambda) {
    ridge <- ridgereg(formula = formula, data = train, lambda = lambda) 
    pred <- predict(ridge, newdata = val)                               
    obs <- val[[all.vars(formula)[1]]]                                  
    RMSE(pred = pred, obs = obs)                                        
  }

  return(objective_function)
}
```

The next step is to loop over a grid of lambda values and select the one with the best (lowest) RMSE.

```{r}
# Create the objective function for ridge regression using the training and validation sets.
# This objective function will return the RMSE for each value of lambda.
objective <- create_ridge_objective_function(
  formula = dep_delay ~ .,   # Model formula: predict dep_delay using all other variables
  train = X_train,           # Training data
  val = X_val                # Validation data
)

# Define a grid of lambda values to test, ranging from 1 to 50 with a step of 1.
lambda_grid <- seq(1, 50, by = 1)

# Initialize an empty vector to store the RMSE results for each lambda value.
results <- c()

# Loop through each lambda value in the grid.
for (lambda in lambda_grid) {
  # Evaluate the model for the current lambda value and store the resulting RMSE.
  results <- c(results, objective(lambda))
}

# Retrieve best lambda
best_lambda <- lambda_grid[which.min(results)]

# Plot the RMSE values as a function of lambda.
plot(lambda_grid, results, type = "b", xlab = "Lambda", ylab = "RMSE", main = "RMSE vs Lambda")

```

The lowest value corresponds to 位 = 32.

## 5. Predict on test set

Once we found a good value for 位, we can use it to predict the test dataset.

```{r}

ridge <- ridgereg(formula = dep_delay ~ ., data = X_train, lambda = best_lambda) 
pred <- predict(ridge, newdata = X_test)
obs <- X_test[["dep_delay"]]                                  
rmse <- RMSE(pred = pred, obs = obs)

print(sprintf("The test RMSE is %s", rmse))
```
