---
title: "linreg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{linreg}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(heleramcar4)
```

|                                                             |
|:------------------------------------------------------------|
| Vignette: A Simple Implementation of Linear Regression in R |

This vignette demonstrates the usage of a simple linear regression
package implemented in R. The package allows users to fit a linear
regression model, get residuals, predictions, and coefficients, as well
as generate plots for model diagnostics.

### Introduction

This vignette demonstrates the usage of a simple linear regression
package implemented in R. The package allows users to fit a linear
regression model, get residuals, predictions, and coefficients, as well
as generate plots for model diagnostics.

### Loading the Package

The package requires `ggplot2` for generating diagnostic plots. Make
sure you have `ggplot2` installed:

```{r install-ggplot2, eval=FALSE}
install.packages("ggplot2")
```

Once the package is installed, you can load it using:

```{r load-library, echo=TRUE, eval=FALSE}
library(ggplot2)
```

### Linear Regression Implementation

The linear regression is implemented using `RefClass` objects, which
allow us to define a class-like structure in R. The package provides the
following main components:

-   `linreg`: The main class that fits the linear regression model.
-   `linregResults`: Stores the results of the regression (coefficients,
    residuals, fitted values, etc.).
-   Several generic functions for accessing residuals, predictions,
    coefficients, and model summaries.

### Creating a Linear Regression Model

To fit a linear regression model, you can use the RC class `linreg`.
This function takes a formula and a data frame as input and returns a
`linreg` object.

Fitting a linear regression model:

```{r fit-model, echo=TRUE}
linreg_mod <- linreg$new(Petal.Length ~ Species, data = iris)
```

### Model Coefficients

To get the coefficients of the fitted model, you can use the `coef()`
function.

```{r get-coefs, echo=TRUE}
coef(linreg_mod)
```

### Model Residuals

The residuals from the model can be accessed using the `residuals()`
function.

```{r get-residuals, echo=TRUE}
residuals(linreg_mod)
```

### Model Predictions

You can get the predicted (fitted) values using the `pred()` function.

```{r get-predictions, echo=TRUE}
pred(linreg_mod)
```

### Model Summary

The `summary()` function returns a detailed summary of the model,
including coefficients, standard errors, t-values, and p-values.

```{r get-summary, echo=TRUE}
summary(linreg_mod)
```
