---
title: "Linreg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(heleramcar4)
```

 ----------------------------------------------------------------------
 Vignette: A Simple Implementation of Linear Regression in R
 ----------------------------------------------------------------------

This vignette demonstrates the usage of a simple linear regression package
implemented in R. The package allows users to fit a linear regression model,
get residuals, predictions, and coefficients, as well as generate plots for
model diagnostics.

### Introduction
This vignette demonstrates the usage of a simple linear regression package implemented in R. 
The package allows users to fit a linear regression model, get residuals, predictions, and coefficients, as well as generate plots for model diagnostics.

### Loading the Package
The package requires `ggplot2` for generating diagnostic plots. Make sure you have `ggplot2` installed:

```{r install-ggplot2, eval=FALSE}
install.packages("ggplot2")
```

Once the package is installed, you can load it using:

```{r load-library, echo=TRUE, eval=FALSE}
library(ggplot2)
```

### Linear Regression Implementation
The linear regression is implemented using `RefClass` objects, which allow us to define a class-like structure in R. 
The package provides the following main components:

- `LinearRegression`: The main class that fits the linear regression model.
- `LinearRegressionResults`: Stores the results of the regression (coefficients, residuals, fitted values, etc.).
- Several generic functions for accessing residuals, predictions, coefficients, and model summaries.

### Creating a Linear Regression Model
To fit a linear regression model, you can use the function `linreg()`. 
This function takes a formula and a data frame as input and returns a `LinearRegression` object.



Fitting a linear regression model:

```{r fit-model, echo=TRUE}
model <- linreg(y ~ x, data)
```

### Model Coefficients
To get the coefficients of the fitted model, you can use the `coef()` function.

```{r get-coefs, echo=TRUE}
coef(model)
```

### Model Residuals
The residuals from the model can be accessed using the `residuals()` function.

```{r get-residuals, echo=TRUE}
residuals(model)
```

### Model Predictions
You can get the predicted (fitted) values using the `pred()` function.

```{r get-predictions, echo=TRUE}
pred(model)
```

### Model Summary
The `summary()` function returns a detailed summary of the model, including coefficients, standard errors, t-values, and p-values.

```{r get-summary, echo=TRUE}
ummary(model)
```

